---
title: "Project Report"
author: "jaime"
date: "6/12/2020"
output: md_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Loading and preprocessing the data

We write code for downloading, creating an archive and placing the file in the directory

```{r}
setwd('C:/Users/David/Downloads/John Hopkins/R/data')
if(!file.exists('data')) dir.create('data')
fileUrl <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'
download.file(fileUrl, destfile = 'C:/Users/David/Downloads/John Hopkins/R/data/repdata_data_activity.zip')
unzip('C:/Users/David/Downloads/John Hopkins/R/data/repdata_data_activity.zip', exdir = 'C:/Users/David/Downloads/John Hopkins/R/data')
activityData <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?

We calculate the mean using aggregate and mean function and ignoring the NA. We also input the data for the histogram

```{r}

totalSteps <- aggregate(steps ~ date, activityData, FUN=sum)


hist(totalSteps$steps,
     main = "Total Steps per Day",
     xlab = "Number of Steps")

meanSteps <- mean(totalSteps$steps, na.rm = TRUE)
medSteps <- median(totalSteps$steps, na.rm = TRUE)

```

## What is the average daily activity pattern?

For the average daly pattern, we plot 5-minute intervals that we previously calculated using the mean. 
We also calculate de max number of steps of any interval

```{r}
library(ggplot2)
meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps") +
  theme(plot.title = element_text(hjust = 0.5))
maxInt <- meanStepsByInt[which.max(meanStepsByInt$steps),]
```

## Imputing Missing Values

We use the function to calculate number of missing values in the dataset. 

```{r}
missingVals <- is.na(activityData$steps)
```

Missing vallues=17568

We then fill the dataset so it doesn't have missing values

```{r}
imp_activityData <- transform(activityData,
                              steps = ifelse(is.na(activityData$steps),
                                             meanStepsByInt$steps[match(activityData$interval, 
                                                                        meanStepsByInt$interval)],
                                             activityData$steps))
```

We now plot the histogram

```{r}
impStepsByInt <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impStepsByInt$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")
```

We calculate the means to calculate the difference we were asked

```{r}
impMeanSteps <- mean(impStepsByInt$steps, na.rm = TRUE)
impMedSteps <- median(impStepsByInt$steps, na.rm = TRUE)
diffMean = impMeanSteps - meanSteps
diffMed = impMedSteps - medSteps
diffTotal = sum(impStepsByInt$steps) - sum(totalSteps$steps)
```

## [1] 10766.19
## [1] 10766.19


Theres 0 difference in the mean of the 2 datasets.

There is a difference of -1.076381110^{4} in the median steps of the two dataset. There is a difference of 8.612950910^{4} in the total steps of the two dataset.

## Are there differences in activity patterns between weekdays and weekends?

We create the new factor with the levels weekday and weekend. We use the dataset without the missing values

```{r}
library(psych)
library(lubridate)
week <- wday(imp_activityData$date)
week_day <- week
for (i in 1:17568) # loop to find the na
{
    if(week[i] == 1)
        week_day[i] <- 'weekend'
    if(week[i] == 2)
        week_day[i] <- 'weekday'
    if(week[i] == 3)
        week_day[i] <- 'weekday'
    if(week[i] == 4)
        week_day[i] <- 'weekday'
    if(week[i] == 5)
        week_day[i] <- 'weekday'
    if(week[i] == 6)
        week_day[i] <- 'weekday'
    if(week[i] == 7)
        week_day[i] <- 'weekend'
}

 
imp_activityData$weekday <-week_day
```

Finding the elements by "weekday" or "weekend"
What  is the average daily activity pattern?

Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, 
averaged across all days (yaxis)

```{r}

weekday <- grep("weekday",imp_activityData$weekday)
weekday_frame <- imp_activityData[weekday,]
weekend_frame <- imp_activityData[-weekday,]

five_minutes_average_weekday <- aggregate(steps~interval, data=weekday_frame, FUN=mean, na.rm=TRUE)
five_minutes_average_weekend <- aggregate(steps~interval, data=weekend_frame, FUN=mean, na.rm=TRUE)

plot(x = five_minutes_average_weekday$interval, y = five_minutes_average_weekday$steps, type = "l")
plot(x = five_minutes_average_weekend$interval, y = five_minutes_average_weekend$steps, type = "l") 
```
"C:/Users/David/Downloads/John Hopkins/R/data/PA1_template.Rmd"
